Synthetic Observations of Extremely High Velocity (EHV) SiO component in young molecular outflows

### 2013-10-29 PAPER/Paola_1/ms_ver2.tex
Final comments of Paola

In-spite of all numerical models relating to the study of enhancement
of SiO in shocks, very little is known about the dependence of SiO
abundance on shock speeds. Considering the complex grain
chemistry that is involved in order to estimate the functional
dependence of SiO abundance on shock speeds, we prescribe these
profiles based on limited empirical evidences. The most simple among
them is the top hat profile in which the SiO abundance is a low value of
10$^{-12}$ below 20 km s$^{-1}$ and above 100 km s$^{-1}$ and a
maximum value of 10$^{-6}$ between these two velocities.
In order to get rid of a discontinuous change of abundance, we also
prescribe a gaussian such that the peak SiO abundance of 10$^{-6}$ is at
60\,km\,s$^{-1}$, while the value of n(SiO)/n(H$_{2}$) falls below
10$^{-9}$ at 20 and 100 km s$^{-1}$
\citep[e.g.,][also see
section~\ref{sec:discussion}]{Lefloch:1998p13983}. 
These functional dependences of SiO
abundance on shock velocity are show in figure~\ref{abun}.  
In addition to this, if the temperature at the point is greater than
92,000\,K (the temperature of the Si-O bond disassociation energy) 
the abundance is reduced to 0. 
%

@@

\section{Conclusion}
\label{sec:conclusion}
We have performed MHD simulations
of episodic radiative jet propagating 
with a typical speed of 100 km s$^{-1}$ into a cold, non-magnetized ambient
molecular medium with a density gradient. The jet dynamical quantities are evolved in conjunction
with different optically thin, non-equilibrium cooling prescriptions of varying complexities. 
The most complex is that of molecular cooling along with
H$_2$ chemistry. This prescription allows us to track the formation
and destruction of HI, HII and H$_2$ along with the flow dynamics. 
The final state of the jet obtained for each cooling model is then
used as input into a non-LTE radiative transfer code to
obtain SiO emission maps, spectra and PV digrams. The main results
obtained from our study are as follows :
\begin{itemize}
\item Different cooling prescriptions used for our model
  significantly influence thickness of the jet. An
  efficient mode of cooling does reduce the size of the cocoon and results
  in a more narrow jet. In addition, the primary bow shock structure
  differs with various cooling modes. Also, with realistic cooling
  functions one sees development of
  thermal instability at the jet head which also shows a significant emission in SiO
  along with young internal knots that are formed close to the base of
  the jet.
\item Emission maps obtained from our reference run with molecular
  cooling show a distinctive outflow lobe like structure. The
  physical regions responsible of SiO emission in such early outflows
  typically have molecular hydrogen densities n$\sim$ 10$^{6}$
  cm$^{-3}$ and temperatures $>$ 500\,K which is consistent with
  properties deduced from LVG modeling of observational data of
  young outflows like L1448-mm.
\item Our multi-line study of SiO clearly shows how different
  line transitions are sensitive to different regions in the jets. For
  example, the SiO J = 1-$>$ 0 transition is mostly excited due to
  the interaction of the jet with the ambient medium, while SiO J = 8-$>$ 7 trace the inner most jet regions as
  they are excited only in the internal knots. This finding very well
  supports the multi-line observational surveys done for young
  molecular outflows, particularly, HH 211.
\item Single-dish synthetic spectra obtained from our model typically
  have intensity of 0.3\,K and peak around 50\,km\,s$^{-1}$ and have a
  typical line width of 20\,km\,s$^{-1}$. Such modeled spectra are in
  excellent agreement with characteristic features of EHV emission of
  SiO seen in young molecular outflows . In addition, the line ratios 
  T(8-7)/T(5-4) and T(5-4)/T(2-1), predicted from our models have values close to
  unity consistent with observations. Further,
  indications of the formation of the slow component due to the
  slowing down of the
  older knots is clearly seen as a shift of centroid velocities by
  7-10\,km\,s$^{-1}$ in the first 10$^{3}$ years of evolution.
\item The predicted PV diagrams from our models show a distinctive
  knot like features that are generally observed in outflows like IRAS
  04166+2706.
\item Emission from internal knots do depend on the chosen profile of
  the abundance. On one extreme where all the internal knots 
  is the top-hat profile with $\delta <$ 1, whereas, no emission in
  seen in knots when a gaussian profile with $\delta$ = 1 is used. For
  the remaining two possibilities, we observe that emission mainly
  comes from knots close to the base of the flow and the condensation
  formed at the primary shock due to thermal instability.
\item Predicted ALMA maps also show characteristic features that are
  consistent with young molecular outflows with EHV emission. 
\end{itemize}

In summary, this work for the first time provides a model
that explains the origin of EHV emission seen in SiO from 
the inner regions of young molecular outflows.

>>>>

In-spite of all numerical models relating to the study of enhancement
of SiO in shocks, very little is known about the dependence of SiO
abundance on shock speeds. Considering the complex grain
chemistry that is involved in order to estimate the functional
dependence of SiO abundance on shock speeds, we prescribe these
profiles based on limited empirical evidences. The most simple among
them is the top hat profile in which the SiO abundance is a low value of
10$^{-12}$ below 20 km s$^{-1}$ and above 100 km s$^{-1}$ and a
maximum value of 10$^{-6}$ between these two velocities.
Shock velocities $>$ 100 km s$^{-1}$ can lead to the 
ionisation of hydrogen and thereby effectively reduce the number
of collision centers for SiO resulting in a weaker emission. 
However, we also consider a step function dependence of SiO abundance
on shock velocity to incorporate cases where we allow for reformation
of atomic and molecular species.
In order to get rid of a discontinuous change of abundance, we also
prescribe a gaussian such that the peak SiO abundance of 10$^{-6}$ is at
60\,km\,s$^{-1}$, while the value of n(SiO)/n(H$_{2}$) falls below
10$^{-9}$ at 20 and 100 km s$^{-1}$. Such a huge variation in SiO
abundance may indeed be the case in regions where SiO is seen
in widespread \citep[e.g.,][also see section~\ref{sec:discussion}]
{Lefloch:1998p13983}. Also along with a Gaussian profile, we consider
the case where we extent the abundance of 10$^{-6}$ beyond 60 km\,s$^{-1}$. 
These functional dependences of SiO
abundance on shock velocity are shown in figure~\ref{abun}.  
In addition to this, if the temperature at the point is greater than
92,000\,K (the temperature of the Si-O bond disassociation energy) 
the abundance is reduced to 0. 
%

@@

\section{Conclusion}
\label{sec:conclusion}
We have performed MHD simulations
of episodic radiative jet propagating 
with a typical speed of 100 km s$^{-1}$ into a cold, non-magnetized ambient
molecular medium with a density gradient. The jet dynamical quantities are evolved in conjunction
with different optically thin, non-equilibrium cooling prescriptions of varying complexities. 
The most complex is that of molecular cooling along with
H$_2$ chemistry. This prescription allows us to track the formation
and destruction of HI, HII and H$_2$ along with the flow dynamics. 
The final state of the jet obtained for each cooling model is then
used as input into a non-LTE radiative transfer code to
obtain SiO emission maps, spectra and PV digrams. The main results
obtained from our study are as follows :
\begin{itemize}
\item Different cooling prescriptions used for our model
  significantly influence thickness of the jet. An
  efficient mode of cooling does reduce the size of the cocoon and results
  in a more narrow jet. In addition, the primary bow shock structure
  differs with various cooling modes. Also, with realistic cooling
  functions one sees development of
  thermal instability at the jet head which also shows a significant emission in SiO
  along with young internal knots that are formed close to the base of
  the jet.
\item Emission maps obtained from our reference run with molecular
  cooling show a distinctive outflow lobe like structure. The
  physical regions responsible of SiO emission in such early outflows
  typically have molecular hydrogen densities n$\sim$ 10$^{6}$
  cm$^{-3}$ and temperatures $>$ 500\,K which is consistent with
  properties deduced from LVG modeling of observational data of
  young outflows like L1448-mm.
\item Our multi-line study of SiO clearly shows how different
  line transitions are sensitive to different regions in the jets. For
  example, the SiO J = 1-$>$ 0 transition is mostly excited due to
  the interaction of the jet with the ambient medium, while SiO J = 8-$>$ 7 trace the inner most jet regions as
  they are excited only in the internal knots. This finding very well
  supports the multi-line observational surveys done for young
  molecular outflows, particularly, HH 211.
\item Single-dish synthetic spectra obtained from our model typically
  have intensity of 0.3\,K and peak around 50\,km\,s$^{-1}$ and have a
  typical line width of 20\,km\,s$^{-1}$. Such modeled spectra are in
  excellent agreement with characteristic features of EHV emission of
  SiO seen in young molecular outflows . In addition, the line ratios 
  T(8-7)/T(5-4) and T(5-4)/T(2-1), predicted from our models have values close to
  unity consistent with observations. Further,
  indications of the formation of the slow component due to the
  slowing down of the
  older knots is clearly seen as a shift of centroid velocities by
  7-10\,km\,s$^{-1}$ in the first 10$^{3}$ years of evolution.
\item The predicted PV diagrams from our models show a distinctive
  knot like features that are generally observed in outflows like IRAS
  04166+2706.
\item Emission from internal knots do depend on the chosen profile of
  the abundance. On one extreme where all the internal knots emit SiO  
  with a step profile with $\delta =$ 1 and a top-hat profile with
  $\delta < 1$, whereas, no emission in
  seen in knots when a gaussian profile with $\delta$ = 1 is used. For
  the remaining cases, we observe that emission mainly
  comes from knots close to the base of the flow and the condensation
  formed at the primary shock due to thermal instability.
\item Predicted ALMA maps also show characteristic features that are
  consistent with young molecular outflows with EHV emission. 
\end{itemize}

In summary, this work for the first time provides a model
that explains the origin of EHV emission seen in SiO from 
the inner regions of young molecular outflows. The unique combination
of axisymmetric jet propogation model with molecular cooling and
non-LTE radiative transfer calculations can sucessfully predict most
of the observed properties of SiO emission from young molecular
outflows from Class 0 stages. This model provides an excellent
template to compare current ALMA obervations and can also predict
signatures of thermal instabilities which can be tested with upcoming
ALMA observations. In future, we would like to extent our
model to full three dimensions and evolve the jet for longer time
scales to explore slow velocity component, precession and 
slew of other instabilities. Additionally, we would also like to have
a better handle on SiO abundance not only from observations but also
by including dust micro-physics in our model.
